Analysis of ToothGrowth data
========================================================


## Synopsis
We're going to analyze the ToothGrowth data in the R datasets package. 

## Basic Summary
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

A data frame with 60 observations on 3 variables.

*  **len**	numeric	Tooth length
*	**supp**	factor	Supplement type (VC or OJ).
*	**dose**	numeric	Dose in milligrams.

## Data analysis
### Tooth growth by supplement
There are two supplements, going by the names VC and OJ. Both of them seem to follow
a normal distribution

```{r}
ToothGrowth.vc <- ToothGrowth[ToothGrowth$supp=="VC",]
hist(ToothGrowth.vc$len,xlab="Tooth length", main= "Supplement VC")
ToothGrowth.oj <- ToothGrowth[ToothGrowth$supp=="OJ",]
hist(ToothGrowth.oj$len, xlab="Tooth length", main= "Supplement OJ")
```
### Toothgrowth by dosage
The same holds true for the dosages:
```{r}
ToothGrowth.lowdose <- ToothGrowth[ToothGrowth$dose==0.5,]
hist(ToothGrowth.lowdose$len, xlab="Tooth length", main= "Dosage: 0.5")
ToothGrowth.middose <- ToothGrowth[ToothGrowth$dose==1.0,]
hist(ToothGrowth.middose$len, xlab="Tooth length", main= "Dosage: 1.0")
ToothGrowth.highdose <- ToothGrowth[ToothGrowth$dose==2.0,]
hist(ToothGrowth.highdose$len, xlab="Tooth length", main= "Dosage: 2.0")
```

## Hypothesis tests for supplement and dosage types
## Supplement
Let us use the two sample T test to determine with conviction the fact that one supplement is more effective than the other.

Basically, let the null hypothesis be that both the dosage types have the same mean.
Alternative hypothesise, hence will be that both the dosage types have unequal means.

We use R's Welch's T-test to determine this.
```{r}
tTest <- t.test(ToothGrowth.oj$len, ToothGrowth.vc$len)
tTest
```
The p-value obtained is `r tTest$p.value` . Since it is greater than the 0.05, on a 95% confidence interval, we cannot reject the null hypothesis. It is however pretty close to that. Depending on the criticality, we can consider it so.

## Dosage

Let's do this pairwise between, 0.5 and 1, and 1 and 2. If we reject the null hypothesis. We can assume transitivity, for all practical purposes.

### Between 0.5mg and 1.0mg
```{r}
tTest <- t.test(ToothGrowth.lowdose$len, ToothGrowth.middose$len)
tTest
```
The p-value obtained is `r tTest$p.value` . Since it is significantly less than the 0.05, on a 95% confidence interval, we can safely reject the null hypothesis.

This essentially means that 1.0 mg, with the tooth length mean as `r mean(ToothGrowth.middose$len)` has a definitively high effectiveness vis-a-vis 0.5 mg (mean: `r mean(ToothGrowth.lowdose$len)`).

### Between 1.0mg and 2.0mg
```{r}
tTest <- t.test(ToothGrowth.middose$len, ToothGrowth.highdose$len)
tTest
```
The p-value obtained is `r tTest$p.value` . Since it is significantly less than the 0.05, on a 95% confidence interval, we can safely reject the null hypothesis.

This essentially means that 1.0 mg, with the tooth length mean as `r mean(ToothGrowth.middose$len)` has a definitively lower effectiveness vis-a-vis 2.0 mg (mean: `r mean(ToothGrowth.highdose$len)`).

## Conclusion
This effectively means that orange juice, with mean `r mean(ToothGrowth.oj$len)` has a higher mean than ascorbic acid, but the p value test proves that it is not conclusive enough for 95% confidence interval. Meanwhile, it has been shown above that greater the dosage, greater is the effect on the tooth length.